Relationships in the extremes and their influence on competition and coexistence

### 2022-03-01 paper/Paper.Rnw
about to remove the lottery model theory from Theory and put it in a new sup mat section

%***DAN: consider sending the lottery model components of this section to sup mat, for length reasons.

@@

%G2
Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. 
First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
it was then sufficient to decompose the invasion growth rate of the weaker competitor, species 1. 
Second, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
against $\mu_1 - \mu_2$. 
Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
and $\mu_1-\mu_2=0,-0.5,-2,-4$,
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
the contributions of other mechanisms.
Mathematical simplications for the $\delta=1$ case make it straightforward to see that
coexistence is impossible and that the case is relatively trivial (SI 
section \ref{SI-sect:delta1}). So that case is not considered further.  
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section \ref{SI-sect:efficient_computation}.

>>>>

%***DAN: consider sending the lottery model components of this section to sup mat, to reduce length and complexity.
%Before doing that, make sure you understand what the consequences will be.

@@

%G2
Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. 
First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
it was then sufficient to decompose the invasion growth rate of the weaker competitor, species 1. 
Second, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
against $\mu_1 - \mu_2$. 
Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
and $\mu_1-\mu_2=0,-0.5,-2,-4$,
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
the contributions of other mechanisms.
Mathematical simplications for the $\delta=1$ case make it straightforward to see that
coexistence is impossible and that the case is relatively trivial (SI 
section \ref{SI-sect:delta1}), so that case is not considered further.  
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section \ref{SI-sect:efficient_computation}.
